Global Life Expectancy: Unraveling Health and Economic Determinants (2015-2020)\thanks{Code and data are available at: \url{https://github.com/wendyhuan/lifeexpectancy}.}

### 2024-12-04 paper/paper.tex
momo

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

@@

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\begin{document}
\maketitle
\begin{abstract}
This paper analyze life expectancy and its determinants through 184
countries and make the prediction of people from different Income Group.
Multiple Linear regression is used to deploying the life expectanc with
gender, income and region. Predictions of life expectancy of people from
different income group is made according to these essential predictors.
The finding indicates that the life expectancy is tend to get higher as
the economic class grows. And we predict that the average age of person
from low, lower-middle, upper-middle and high are 65,70, 75,80. The
result of this paper will help in suggesting a country which area should
be given importance in order to efficiently improve the life expectancy
of its population.
\end{abstract}

@@

Life expectancy is a critical measure of a population's overall health
and well-being, shaped by various factors such as gender, geographic
location and economic group. The index of life expectancy is generally
served as a benchmark for income group, indicating the effectiveness of
interventions in reducing mortality and improving well-being. Higher
life expectancy is believed to linked to stronger physical factor,
better living standards, and geographic location. Conversely, low life
expectancy often signals systemic challenges such as weaker health,
lower economic situation, and inadequate healthcare access of different
region. (\textbf{socialdeterminantsofhealth?}).

@@

Related Research has shown that there exist difference between men and
women related with biological, behavioral, and socioeconomic factors,
highlighting that gender-specific health behaviors and societal roles
influence longevity (\textbf{GenderandLifeExpectancy?}). Higher-income
individuals tend to live longer due to better access to healthcare and
healthier lifestyles, with notable regional disparities even within
similar income groups (\textbf{IncomeGroupsandLongevity?}).
Additionally, regional socioeconomic differences on health outcomes,
especially life expectancy. It emphasizes that addressing regional
inequities in wealth and resources is essential for improving population
health globally. (\textbf{RegionalVariations?}).

@@

The structure of the paper is as follows: Section~\ref{sec-data}
outlines the data sources and variables considered, followed by the
model setup in Section~\ref{sec-modset} and justification in
\textbf{?@sec-modjust}. The results in \textbf{?@sec-result} presents
the key findings of the analysis, with a discussion on the implications.
\textbf{?@sec-discussion} then discusses potential limitations and
suggestions for future research. \textbf{?@sec-appx} provides additional
detailed information about the data, model and methodology.

@@

The data used in this analysis originates from The World Health
Organization's (WHO) and Global Health Observatory (GHO)
(\textbf{lifeexpectancy?}). This data-set related to life expectancy,
health factors for 184 countries has been collected from WHO data
repository website and its corresponding economic data was collected
from United Nation website. Among all categories of health-related
factors only those socio-economic factors on the national level were
chosen for global scale analysis.

This analysis uses the statistical programming language R (R Core Team
2023) and several libraries, including \texttt{tidyverse}
(\textbf{tidyverse?}), \texttt{janitor} (\textbf{janitor?}),
\texttt{knitr} (\textbf{knitr?}), \texttt{dplyr} (\textbf{dplyr?}),
\texttt{arrow} (\textbf{arrow?}), \texttt{purrr} (\textbf{purrr?}),
\texttt{sf} (\textbf{sf?}), and \texttt{here} (\textbf{here?}) for data
manipulation. \texttt{ggplot2} (\textbf{ggplot?}), \texttt{ggcorrplot}
(\textbf{ggcorrplot?}) and \texttt{kableExtra} (\textbf{kableExtra?})
for visualization. The dataset covers various predictors conducted
across multiple countries, capturing the support for a country to
determine the predicting factor which is contributing to lower value of
life expectancy.

@@

\textbf{Region(Region)}: This is a categorical variable that classifies
the geographic location of the countries. It is mapped one by one by the
name of the country. The data is stored as the name of the
continent(`Africa', `Oceania', `Asia', `Europe', `North America', `South
America').

@@

To make such table, we first select and rename key variables from raw
data to focus on relevant information. To make the subsequent analysis
easier, we then convert variables to the proper data types and eliminate
the rows that have missing data values. To keep things neat, we organize
the decimal for every piece of numerical data and drop the percentage
symbol. The income group column and the region column was not given in
the raw dataset. We created a mapping over country name to four income
group according to the index given by World Bank Income Group and save
the data under ``Income\_Group''. We as well made a mapping to the
countries according to its continent and saved as ``Region''. We again
merge the table, removing repeated columns and rows. For easier
visualization, we created four tables, which are grouped by life
expectancy at age 60, life expectancy at birth, life expectancy at birth
of Male and life expectancy at birth of Female.

shows the average of life expectancy under male, female, 6 different
regions and 4 income groups.

@@

More information on the data cleaning process can be found in
\textbf{?@sec-appx}.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Income Group(\texttt{Income\_Group})}: The income group of a
  country is the key factor influencing the life expectancy. This
  variable provides context for comparing life expectancy between
  different levels of country income group. The division of the life
  expectancy by the income group of a country because income levels
  often correlate strongly with various factors affecting health and
  longevity. Higher income countries typically have more resources to
  invest in robust healthcare systems, with better living conditions
  while lower income countries often face challenges such as inadequate
  healthcare infrastructure and malnutrition.
\end{enumerate}

Following by the four income group, we would like to see the
distribution of the income group of 184 WHO member countries.
(\textbf{fid-income?}) shows that there are 55 countries at the high
income group, 26 countries at low income group, 54 countries at the
lower-middle group while there are 49 countries at the upper-middle
group. With significant less low income countries, we expect to see the
life expectancy lie in a comparably higher range. On the other hand,
there might exist sever outliers dropping the mean.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\item
  \textbf{Gender(\texttt{Gender})}: The gender of the population of each
  country is included as a categorical variable (Male/Female/Both Sex).
  This is a key variable in the dataset because gender has been fully
  recognized connected with the biological physical factor which
  directly effect the life expectancy. Figure~\ref{fig-gender} shows
  that the with an stable average of 74, life expectancy of female is
  constantly higher than male's average of 70. The line plot reveals a
  steady increase until 2019, followed by a sharp decline in 2020, which
  is assumed to be associated with the impact of COVID-19.
\item
  \textbf{Region(Region)}: This is a categorical variable that
  classifies the geographic location of the countries. It is mapped one
  by one by the name of the country. The data is stored as the name of
  the continent(`Africa', `Oceania', `Asia', `Europe', `North America',
  `South America'). Figure~\ref{fig-region} indicates that with the
  number of 54, Africa has the most WHO member countries. Both Asia and
  Europe has 43 countries,listing in the middle while the other three
  continent has the significant less number of countries. North America
  has 22 countries, South America has 12 countries and Oceania has the
  least, 10 countries.
\end{enumerate}

@@

\caption{\label{fig-income}Distribution of country income group.}

@@

\begin{table}
\centering
\caption{Summary Statistics of Life Expectancy}
\centering
\begin{tabular}[t]{c|c|c|c|c|c}
\hline
Mean & Median & Max & Standard Deviation & Variance & N\\
\hline
45.9 & 37.7 & 87.4 & 26.9 & 724.2 & 6624\\
\hline
\end{tabular}
\end{table}

@@

\caption{\label{fig-summary}Summary statistics of the number of life
expectancy over countries and years.}

@@

\begin{align} 
y_i|\mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_1 \text{Region}_i + \beta_2 \text{Gender}_i + \beta_3 \text{Income_Group}_i \\
\alpha &\sim \text{Normal}(0, 2.5) \\
\beta_1, \beta_2, \beta_3 &\sim \text{Normal}(0, 2.5) \\
\sigma &\sim \text{Exponential}(1)
\end{align}

@@

We run the model in R (R Core Team 2023) using the \texttt{rstanarm}
package of Goodrich et al. (2022). We use the default priors from
\texttt{rstanarm}. More explanation of the model could be found
Appendix~\ref{sec-model-details}.

\subsubsection{Model justification}\label{model-justification}

@@

\paragraph{Model Validation}\label{model-validation}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#model\_developed \textless{}{-}}
  \CommentTok{\#readRDS(file = "\textasciitilde{}/lifeexpectancy/models/ledata\_normal\_model.rds")}

\CommentTok{\# Split the dataset into training and testing sets (80\% train, 20\% test)}
\CommentTok{\#set.seed(123)  \# Setting seed for reproducibility}
\CommentTok{\#train\_index \textless{}{-} createDataPartition(leab$\textasciigrave{}Life Expectancy\textasciigrave{}, p = 0.8, list = FALSE)}

\CommentTok{\#train\_data \textless{}{-} leab[train\_index, ]}
\CommentTok{\#test\_data \textless{}{-} leab[{-}train\_index, ]}

\CommentTok{\# Predict life expectancy using the test data}
\CommentTok{\#predictions \textless{}{-} predict(ledata\_normal\_model, newdata = test\_data)}
\end{Highlighting}
\end{Shaded}

\section{Results}\label{results}

@@

Our results are summarized in Table~\ref{tbl-modelresults}. We are
primarily interest in the

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[landscape]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

@@

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{siunitx}

    \newcolumntype{d}{S[
      table-align-text-before=false,
      table-align-text-after=false,
      input-symbols={-,\*+()}
    ]}

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\begin{document}
\maketitle
\begin{abstract}
This paper analyze life expectancy and its determinants through 184
countries and make the prediction of people from different Income Group.
Multiple Linear regression is used to deploying the life expectanc with
gender, income and region. Predictions of life expectancy of people from
different income group is made according to these essential predictors.
The finding indicates that the life expectancy is tend to get higher as
the economic class grows. And we predict that the average age of person
from low, lower-middle, upper-middle and high are 62.35, 68.40, 73.38,
79.38. The result of this paper will help in suggesting a country which
area should be given importance in order to efficiently improve the life
expectancy of its population.
\end{abstract}

@@

Life expectancy is a critical measure of a population's overall health
and well-being, shaped by various factors such as gender, geographic
location and economic group. The index of life expectancy is generally
served as a benchmark for income group, indicating the effectiveness of
interventions in reducing mortality and improving well-being. Higher
life expectancy is believed to linked to stronger physical factor,
better living standards, and geographic location. Conversely, low life
expectancy often signals systemic challenges such as weaker health,
lower economic situation, and inadequate healthcare access of different
region. (Wilkinson 2003).

@@

Related Research has shown that there exist difference between men and
women related with biological, behavioral, and socioeconomic factors,
highlighting that gender-specific health behaviors and societal roles
influence longevity (Oksuzyan 2008b). Higher-income individuals tend to
live longer due to better access to healthcare and healthier lifestyles,
with notable regional disparities even within similar income groups
(Chetty 2016). Additionally, regional socioeconomic differences on
health outcomes, especially life expectancy. It emphasizes that
addressing regional inequities in wealth and resources is essential for
improving population health globally. (F. Marmot M. 2008).

@@

The structure of the paper is as follows: Section~\ref{sec-data}
outlines the data sources and variables considered, followed by the
model setup in Section~\ref{sec-modset} and justification in
Section~\ref{sec-modjust}. The results in Section~\ref{sec-result}
presents the key findings of the analysis, with a discussion on the
implications. Section~\ref{sec-discussion} then discusses potential
limitations and suggestions for future research. Section~\ref{sec-appx}
provides additional detailed information about the data, model and
methodology.

@@

The data used in this analysis originates from The World Health
Organization's (WHO) and Global Health Observatory (GHO) (WHO 2020).
This data-set related to life expectancy, health factors for 184
countries has been collected from WHO data repository website and its
corresponding economic data was collected from United Nation website.
Among all categories of health-related factors only those socio-economic
factors on the national level were chosen for global scale analysis.

This analysis uses the statistical programming language R (R Core Team
2023) and several libraries, including \texttt{tidyverse} (Wickham et
al. 2019), \texttt{janitor} (Firke 2024), \texttt{knitr} (Xie 2024),
\texttt{dplyr} (Wickham, François, et al. 2023), \texttt{arrow}
(Richardson et al. 2023), \texttt{purrr} (Wickham and Henry 2023),
\texttt{sf} (Pebesma and Bivand 2023), and \texttt{here} (Müller 2023)
for data manipulation. \texttt{ggplot2} (Wickham, Chang, et al. 2023),
\texttt{ggcorrplot} (Kassambara 2029) and \texttt{kableExtra} (Zhu 2023)
for visualization. The dataset covers various predictors conducted
across multiple countries, capturing the support for a country to
determine the predicting factor which is contributing to lower value of
life expectancy.

@@

\textbf{Region(\texttt{Region})}: This is a categorical variable that
classifies the geographic location of the countries. It is mapped one by
one by the name of the country. The data is stored as the name of the
continent(`Africa', `Oceania', `Asia', `Europe', `North America', `South
America').

@@

To make such table, we first select and rename key variables from raw
data to focus on relevant information. To make the subsequent analysis
easier, we then convert variables to the proper data types and eliminate
the rows that have missing data values. To keep things neat, we organize
the decimal for every piece of numerical data and drop the percentage
symbol. The income group column and the region column was not given in
the raw dataset. We created a mapping over country name to four income
group according to the index given by World Bank Income Group and save
the data under ``Income\_Group''. We as well made a mapping to the
countries according to its continent and saved as ``Region''. We again
merge the table, removing repeated columns and rows. For easier
visualization, we created four tables, which are grouped by life
expectancy at age 60, life expectancy at birth, life expectancy at birth
of Male and life expectancy at birth of Female. For easier summary and
convenience for graph drawing, we calculate the average of life
expectancy of each country under 6 years. Also, we calculate the average
of 6 years life expectancy according to the other 4 income groups and
different gender by mutate according to the different factors. We then
merge the columns of each table into a summary table for easy looking.

Table~\ref{tbl-sum_sta} shows the average of life expectancy under male,
female, 6 different regions and 4 income groups. Life expectancy
according to different countries have attached and be found in data
cleaning full data in Section~\ref{sec-appx}.

@@

More information on the data cleaning process can be found in
Section~\ref{sec-appx}.

\begingroup\fontsize{10}{12}\selectfont

\begin{longtable}[t]{lc}

\caption{\label{tbl-sum_sta}This table shows the cleaned data of life
expectancy at different predictor. We could visual that high income has
the highest life expectancy of 79.8, compared to lower income has the
least life expectancy of 62.35. Among all the region, it's clear that
Africa has the least 64.17, while Europe has the highest 78.50 years
old. Female has slightly higher life expectancy than male.}

\toprule
Region/Income/Gender & Life Expectancy\\
\midrule
Global & 72.13\\
Male & 69.71\\
Female & 74.62\\
Africa & 64.17\\
Asia & 73.97\\
\addlinespace
North America & 75.04\\
South America & 75.18\\
Oceania & 70.17\\
Europe & 78.50\\
lower\_income & 62.35\\
\addlinespace
lower\_middle & 68.40\\
upper\_middle & 73.38\\
high\_income & 79.38\\
\bottomrule

\endgroup{}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Income Group(\texttt{Income\_Group})}: The income group of a
  country is the key factor influencing the life expectancy. This
  variable provides context for comparing life expectancy between
  different levels of country income group. The division of the life
  expectancy by the income group of a country because income levels
  often correlate strongly with various factors affecting health and
  longevity. Higher income countries typically have more resources to
  invest in robust healthcare systems, with better living conditions
  while lower income countries often face challenges such as inadequate
  healthcare infrastructure and malnutrition. Following by the four
  income group, we would like to see the distribution of the income
  group of 184 WHO member countries. Figure~\ref{fig-income} shows that
  there are 55 countries at the high income group, 26 countries at low
  income group, 54 countries at the lower-middle group while there are
  49 countries at the upper-middle group. With significant less low
  income countries, we expect to see the life expectancy lie in a
  comparably higher range. On the other hand, there might exist sever
  outliers dropping the mean.
\item
  \textbf{Gender(\texttt{Gender})}: The gender of the population of each
  country is included as a categorical variable (Male/Female/Both Sex).
  This is a key variable in the dataset because gender has been fully
  recognized connected with the biological physical factor which
  directly effect the life expectancy. Figure~\ref{fig-gender} shows
  that the with an stable average of 74, life expectancy of female is
  constantly higher than male's average of 70. The line plot reveals a
  steady increase until 2019, followed by a sharp decline in 2020, which
  is assumed to be associated with the impact of COVID-19.
\item
  \textbf{Region(Region)}: This is a categorical variable that
  classifies the geographic location of the countries. It is mapped one
  by one by the name of the country. The data is stored as the name of
  the continent(`Africa', `Oceania', `Asia', `Europe', `North America',
  `South America'). Figure~\ref{fig-region} indicates that with the
  number of 54, Africa has the most WHO member countries. Both Asia and
  Europe has 43 countries,listing in the middle while the other three
  continent has the significant less number of countries. North America
  has 22 countries, South America has 12 countries and Oceania has the
  least, 10 countries.
\end{enumerate}

@@

\caption{\label{fig-income}Distribution of country income group. Among
184 WHO member countries, there are 55 high income countries, 54 lower
middle income countries and 49 upper middle income countries. There are
significant less lower income countries, which we could assume there
might exist less outliers in life experctanncy.}

@@

\begingroup\fontsize{14}{16}\selectfont

\begin{longtable*}[t]{cccccc}
\toprule
Mean & Median & Max & Standard Deviation & Variance & N\\
\midrule
72.2 & 73.1 & 87.4 & 7.8 & 60.1 & 3312\\
\bottomrule
\end{longtable*}
\endgroup{}

@@

\caption{\label{fig-summary}Summary statistics of the number of life
expectancy over countries and years. Less difference between mean and
median, showing that the life expectancy distribution might shows a
naturally bell curve.However, variance is slightly high indicates that
the tails on both sides might be long.}

@@

\begin{align*} 
y_i|\mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_1 \text{Region}_i + \beta_2 \text{Gender}_i + \beta_3 \text{Income\_Group}_i \\
\alpha &\sim \text{Normal}(0, 2.5) \\
\beta_1, \beta_2, \beta_3 &\sim \text{Normal}(0, 2.5) \\
\sigma &\sim \text{Exponential}
\end{align*}

@@

We run the model in R (R Core Team 2023) using the \texttt{rstanarm}
package of Goodrich et al. (2020). We use the default priors from
\texttt{rstanarm}. More explanation of the model could be found
Appendix~\ref{sec-model-details}.

\subsection{Model justification}\label{sec-modjust}

@@

\subsection{Model Validation}\label{model-validation}

@@

Extended versions of the table can be found in
Table~\ref{tbl-testpredicted}.

\begingroup\fontsize{14}{16}\selectfont

\begin{longtable}[t]{cc}

\caption{\label{tbl-modelpredict}Summary of the life expectancy model,
which includes gender, region and income group. The table presents the
model RSME and out-of sample testing. Seeing the testing data slightly
less than the training data, our model is under good generalization.}

\toprule
Dataset & RMSE\\
\midrule
Test Data & 1.15\\
Training Data & 1.17\\
\bottomrule

\endgroup{}

\section{Results}\label{sec-result}

\subsection{Data results analysis}\label{data-results-analysis}

According to Figure~\ref{fig-summary}, we could tell that the mean life
expectancy over the 6 years is 72.2, with a relatively low standard
deviation of 7.8. This suggests that, on average, life expectancy across
the dataset is fairly consistent, with only moderate variation from the
mean. However, the maximum life expectancy of 87.4 stands out
significantly, indicating that some countries or regions are achieving
substantially higher life expectancy.With the mean at 72.2 and the
median at 73.1, it's apparent that the data is fairly symmetric, with
the median being slightly higher than the mean, suggesting a mild left
skew. This shift, combined with the large maximum value, hints at the
presence of outliers---countries with exceptionally high life expectancy
that pull the maximum up. The variance of 60.1 is relatively large,
further supporting the idea that the data spans a wide range, with some
observations considerably deviating from the central tendency.

Given this, the high variance can be attributed to the diversity in life
expectancy values across different regions, particularly when countries
at the higher end of the spectrum (e.g., those with life expectancy
around 87 years) are compared with those at the lower end. This
variability suggests that while most countries have life expectancy near
72 years, the outliers significantly influence the spread of the data,
leading to a higher variance than what would be expected in a more
homogenous dataset. Therefore, while the standard deviation is
relatively small, the data is far from uniform, and a deeper analysis of
the outliers could provide further insights into the factors
contributing to the high life expectancy in certain regions.

Table~\ref{tbl-sum_sta} summarizes global life expectancy across various
categories such as gender, geographic regions, and income levels. We
could tell that the global life expectancy is 72.13 years, providing a
benchmark against which other categories can be compared. Women have a
significantly higher life expectancy (74.62 years) than men (69.71
years), which reflects a common global trend due to biological and
social factors. Region variation does happens to influence the life
expectncy. Europe has the highest life expectancy (78.50 years), which
is likely due to better healthcare and living conditions, with having
the most amount of high income countries. Africa has the lowest (64.17
years), potentially reflecting challenges like limited healthcare
infrastructure, poverty, and infectious diseases. Income level, as one
of the most essential factor, significantly affect the life expectancy.
High-income countries have the highest life expectancy (79.38 years),
illustrating the correlation between economic wealth and longevity. High
income level are believed to enable better living standard compared to
other income groups. Upper middle income countries has an average of
73.38 years, where we could observe that the difference is significant
lower than high income group. Lower middle income group has the life
expectancy drop to 68.40. While these countries usually face challenges
such as infrastructure and limit access to health resources, economic
growth may need to be taken seriously be policy makers. Lower-income
countries have a significantly lower life expectancy of 62.35 years,
which can be attributed to limited access to healthcare, education, and
basic services. Table~\ref{tbl-sum_sta} highlights global inequalities
and emphasizes the importance of improving living standards, especially
in lower-income and underdeveloped regions.

@@

Our results are summarized in Table~\ref{tbl-model}. We are primarily
interest in the general life expectancy regarding to the related
predictors. The multiple linear model provides us with the estimates for
the intercept and coefficients for the predictors which are male,
female, both genders, lower income, lower-middle, upper-middle, high
income, Asia, Europe, Africa, Oceania, North America and South America.
The intercept represents the estimate of region-standard life expectancy
when all the predictors are zero. The coefficients represent the
additional life expectancy associated with each predictor. The model
results will display the estimates- posterior means or medians for each
coefficient including the intercept, uncertainty measures- credible
intervals. The output values of each predictor is the regression
coefficient, meaning how much the outcome is expected to increase or
decrease with one unit increase in the life expectancy (per year) of
that predictor, holding all else constant. The value in the brackets
represent the Median Absolute Deviation of the posterior distributions
of the coefficients. It conveys the dispersion around the median of each
coefficients' posterior distribution, exhibiting how spread the
distributions are.

Num.Obs represents the number of observations made in the model. R2 is
the R-squared value which is the proportion of variance in the dependent
variable that can be explained bu the independent variable. The R2 adj
is the adjusted R squared which accounts for the number of predictors
used. Log.lik is the log-likelihood which gives us an idea of the
likelihood of the data, higher is the better, but this is typically used
for comparison between the models. ELPD and ELPD s.e. explains the log
predictive density and its standard error. The ELPD measures the sum of
the log predictive densities for each observation, used for model
comparison. LOOIC is an acronym for leave-one-out information criterion
in which a lower value indicates a model with better out-of-sample
predictive performance. WAIC stands for Watanabe-Akaike information
criterion which is another measure of good fit; lower values are better
fit. RMSE is the room mean squared error measuring the models's
predictive performance where lower values means more accurate predicts.

\begin{table}

\caption{\label{tbl-model}Summary of the life expectancy model, which
includes gender, region and income group. The table presents the model
RSME and out-of sample testing.}

\centering\centering
\fontsize{7.5}{9.5}\selectfont
\begin{tabular}[t]{lc}
\toprule
